\chapter{Conclusions \& Future Work}
\label{chap:conc}

\section{Conclusions}

It is an exciting time in gravitational-wave astrophysics and numerical relativity. With the first detection of gravitational waves from a a binary neutron star and the coincident detection of electromagnetic counterparts across the EM spectrum, a lot can be discovered about these extreme systems. With many more detections on the horizon, we will need to be prepared on the theoretical side. To help with the ongoing theoretical effort of properly modelling binary compact object systems and their emissions, this thesis has attempted to make progress along two different directions.

Firstly, we have sought to improve the numerical techniques used to solve the partial differential equations that arise in binary neutron star simulations. To this effect, Chapter 2 presented a completely new scheme based on discontinuous Galerkin methods, and we tested it to solve for constant density star initial data, which contains phase transitions similar to that in neutron stars. We also tested the code on multi-black hole puncture initial data, which contains multiple non-smooth points. All tests showed very promising results. Using a combination of hp-refinement and a multigrid preconditioned additive Schwarz solver, the code converges in very few iterations. The code scales well to 1000s of cores when the grid is uniform. The biggest weakness is parallel scalability in the general hp-refined case. No efforts have been made to optimize for scalability in the present implementation, so this isn't surprising. Currently, work is being done to extend this code by Nils Fischer and Prof. Harald Pfeiffer at the Albert Einstein Institute (AEI) in Potsdam, Germany by porting it into the task based parallel code SpECTRE (\citet*{kidder2016spectre}), which is currently being developed at Cornell, Caltech and AEI to solve hyperbolic problems in numerical relativity on future exascale supercomputers using a task-based parallelism framework. The final goal will be to run our implementation on binary neutron star initial data with realistic microphysics and solve down to accuracies previously unobtainable.

Secondly, we wanted to increase the knowledge-base surrounding emissions from BNS mergers and their connection to binary parameters. Chapter 3 presented twelve state-of-art general relativistic radiation-hydrodynamics simulations of binary neutron star mergers with varying realistic EOS and mass-ratios.  With these simulations we established that previous results using different codes were qualitatively correct, i.e. that the softest models produce the most ejecta and neutrino emission, the prompt collapse models produce little to no ejecta, the $Y_e$ and $v_\infty$ distributions are broad across EOS and mass-ratio, larger disk masses with mass-ratio and stiffness of EOS. This agreement is remarkable when we take into consideration that all of the previous results use different numerical implementations. Furthermore, since we used different masses and EOS for our binaries, our work adds to the growing effort to probe the possible parameter space of BNS mergers and its relation to observables such as matter and neutrino emission.  In the future, our work will be extended in the following ways. Firstly, the subset of simulations which collapsed to a black hole were not evolved past this point and it would be interesting to see what emissions arose afterwards and to study the properties and evolution of the resulting accretion disks, which have been shown to produce a significant amount of ejecta when MHD is introduced (\cite{fernandez2019long}). Secondly, we initially ran each simulation with tracer particles that followed the fluid flow, but stopped this because it was slowing down the code too much to get a report out on time for this thesis. Tracer particles would allow us to better study the properties of the ejecta, by not only tracing out the fluid worldlines of the ejecta, but also by using the tracer data to seed nuclear reaction networks and light curve models. All of this was done for neutron star - black-hole binaries in (\cite{fernandez2016dynamics}) by some of our collaborators. In terms of other source parameters, we also neglected in our study the eccentricity of the binary and the spins of the neutron stars, which could potentially produce significantly more ejecta. In the future, the addition of magneto-hydrodynamics, eccentricity, spins and more realistic neutrino transport schemes will definitely be on the agenda.

In conclusion, while this thesis has helped push us closer to the goal of simulating and understanding binary neutron star mergers, a great deal of work still needs to be done. The next few decades should be a very interesting period for both numerical relativity and gravitational wave astrophysics, so stay tuned.

%% \section{Future Work and Directions}

%% Moving forward, the work in this thesis will spawn new research projects. Firstly, 
%% our discontinuous Galerkin solver will be incorporated into the numerical relativity code SpECTRE (\cite{kidder2016spectre}), which is using a task-based parallelism framwork to prepare for running on exa-scale supercomputers. Secondly, our binary neutron star simulations can be continued to long times post-merger, at which point they can be used to initialize axisymmetric simulations for late-time ejecta analysis such as in (NUCLEOSYNTHESIS paper here). Furthermore, 
